Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-04-05 bbbq_article.tex
Spellcheck, progress #212

In our previous work, we found that the observed pattern can be
explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
Here we do a similar analysis, by correlating the percentage of TMH-derived
epitopes presented versus the mean hydrophobicity of all presented
peptides. From these points, we derive a linear trendline and measure
the proportion of the correlation explained by the assumed linear relation.

@@

To determine if epitopes derived from TMHs are presented \emph{in vivo}
we used two elution studies, both of which have sequenced 
the epitopes found in naturally occurring human immune cells,
for MHC-I (\cite{schellens2015comprehensive}) 
and MHC-II (\cite{bergseng2015different}).
For each of the sequenced epitopes,
we found its possible location(s) in 
a human representative reference proteome,
with UniProt ID UP000005640\_9606.
For the epitopes that were present in the proteome exectly once,
we predicted the topology of the proteins in which these epitopes are found.
We used both TMHMM (\cite{krogh2001predicting}) and PureseqTM (\cite{wang2019efficient}) to do these predictions. 
From this topology, we could determine if the epitope
overlapped with a transmembrane region.

@@

  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
    \label{fig:hydrophobicity_1}
  \end{subfigure}  
  \hfill
  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
    \label{fig:hydrophobicity_2}
  \end{subfigure}  
  \caption{
    \textbf{(A)} 
    TMH-derived peptides are predicted to be over-presented in general,
    which is for MHC-I (left column) and MHC-II (right column),
    as well as for the proteomes of SARS-CoV-2 (top row), human (middle 
    row) and \emph{M. tuberculosis} (bottom row).
    The bars show the percentage of epitopes predicted to be 
    overlapping with TMHs.
    The red lines indicate the percentage as expected by chance.
    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
    for the exact TMH and  epitope counts.
    \textbf{(B)} and \textbf{(C)}:
    The relation between the percentage of TMH-derived epitopes presented,
    and the hydrophobicity score of binders for MHC-I haplotypes (panel B)
    and MHC-II haplotypes (panel C). The trendline assumes a linear relationship.
  }
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hydrophocity explains the generality of this finding}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

In our previous work, we found that the observed pattern can be
explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
Here we do a similar analysis, by correlating the percentage of TMH-derived
epitopes presented versus the mean hydrophobicity of all presented
peptides. From these points, we derive a linear trend line and measure
the proportion of the correlation explained by the assumed linear relation.

@@

To determine if epitopes derived from TMHs are presented \emph{in vivo}
we used two elution studies, both of which have sequenced 
the epitopes found in naturally occurring human immune cells,
for MHC-I (\cite{schellens2015comprehensive}) 
and MHC-II (\cite{bergseng2015different}).
For each of the sequenced epitopes,
we found its possible location(s) in 
a human representative reference proteome,
with UniProt ID UP000005640\_9606.
For the epitopes that were present in the proteome exactly once,
we predicted the topology of the proteins in which these epitopes are found.
We used both TMHMM (\cite{krogh2001predicting}) and PureseqTM (\cite{wang2019efficient}) to do these predictions. 
From this topology, we could determine if the epitope
overlapped with a transmembrane region.

@@

  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
    \label{fig:hydrophobicity_1}
  \end{subfigure}  
  \hfill
  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
    \label{fig:hydrophobicity_2}
  \end{subfigure}  
  \caption{
    \textbf{(A)} 
    TMH-derived peptides are predicted to be over-presented in general,
    which is for MHC-I (left column) and MHC-II (right column),
    as well as for the proteomes of SARS-CoV-2 (top row), human (middle 
    row) and \emph{M. tuberculosis} (bottom row).
    The bars show the percentage of epitopes predicted to be 
    overlapping with TMHs.
    The red lines indicate the percentage as expected by chance.
    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
    for the exact TMH and  epitope counts.
    \textbf{(B)} and \textbf{(C)}:
    The relation between the percentage of TMH-derived epitopes presented,
    and the hydrophobicity score of binders for MHC-I haplotypes (panel B)
    and MHC-II haplotypes (panel C). The trend line assumes a linear relationship.
  }
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hydrophobicity explains the generality of this finding}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
